GLOBALIZER: A NOVEL SUPERCOMPUTER SOFTWARE SYSTEM FOR SOLVING TIME-CONSUMING GLOBAL OPTIMIZATION PROBLEMS

### 2016-08-04 2016/GOM/paper.tex
Update GOM paper

@@

\subsection{Core Multidimensional Generalized Algorithm of Global Search}
\label{subsec:corepar}
The information-statistical theory of global search formulated in (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000)
has served as a basis for the development of a large number of efficient multiextremal
optimization methods --- see, for example, (Gergel, V. P., 1996, 1997), (Gergel, V. P., Strongin, R. G., 2003, 2005), (Grishagin, V. A., Strongin, R. G., 1984), (Sergeyev Y. D., 1995, 1999), (Sergeyev, Y. D., Grishagin V. A., 1994, 2001), (Sergeyev, Y. D., Strongin, R. G., Lera, D., 2013), (Barkalov K. A., Gergel V. P., 2014), etc.
\par
Multidimensional Algorithm of Global Search (MAGS) established the basis for the
methods applied in Globalizer. The general computational scheme of MAGS can be
presented as follows --- see (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000).
\par
Let us introduce a simpler notation for the problem being solved on a computing node
\begin{equation}
\label{eq:oneDimTask}
f(x) = \varphi(y(x)):x\in [0;1]\}
\end{equation}
\par
The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
Let us assume further \(k\), \(k>1\), iterations of a global search are to be completed.
The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}

@@

\section{Globalizer System Architecture}
\label{sec:arch}
\section{Results of Numerical Experiments}
\label{sec:experiments}
\subsection{Parallel Computations Using Intel Xeon Phi processors}
\subsection{Parallel Computations Using General Purpose Graphic Processors}

>>>>

\label{sec:intro}

@@

\subsection{Core Multidimensional Generalized Algorithm of Global Search}
\label{subsec:corepar}
The information-statistical theory of global search formulated in (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000)
has served as a basis for the development of a large number of efficient multiextremal
optimization methods --- see, for example, (Gergel, V. P., 1996, 1997), (Gergel, V. P., Strongin, R. G., 2003, 2005), (Grishagin, V. A., Strongin, R. G., 1984), (Sergeyev Y. D., 1995, 1999), (Sergeyev, Y. D., Grishagin V. A., 1994, 2001), (Sergeyev, Y. D., Strongin, R. G., Lera, D., 2013), (Barkalov K. A., Gergel V. P., 2014), etc.
\par
Multidimensional Algorithm of Global Search (MAGS) established the basis for the
methods applied in Globalizer. The general computational scheme of MAGS can be
presented as follows --- see (Strongin, R. G., 1978), (Strongin, R. G., Sergeyev, Y. D., 2000).
\par
Let us introduce a simpler notation for the problem being solved on a computing node
\begin{equation}
\label{eq:oneDimFunc}
f(x) = \varphi(y(x)):x\in [0,1]\}
\end{equation}
\par
The initial iteration of the algorithm is performed at an arbitrary point \(x^1(0,1)\).
Let us assume further \(k\), \(k>1\), iterations of a global search are to be completed.
The selection of the trial point \(k+1\) for the next iteration is performed according to the following rules.
\subsection{Parallel Computations for Systems with Shared Memory}
\label{subsec:sharedpar}

@@

\section{Globalizer System Architecture}
\label{sec:arch}
As previously mentioned in Section \ref{sec:intro}, the development of software systems
for global optimization is complex professional work. The optimization systems should
include advanced optimization methods, provide for the accumulation and efficient
processing of search information, include various forms of visualization for the results,
provide tools for interacting with users, etc. All the features pointed out above
become much more complicated when organizing parallel computations using high-performance
multiprocessor systems of varying architectures.
\par
The Globalizer considered in this paper expands the family of global optimization
software systems successively developed by the authors during the past several years.
One of the first developments was the SYMOP multiextremal optimization system (see Gergel, V. P., 1993),
which has been successfully applied for solving many optimization problems. A special
place is occupied by the ExaMin system (see, for example, (Barkalov, K., Gergel V., 2015)),
which was developed and used extensively to investigate the application of novel parallel
algorithms to solve global optimization problems using high-performance multiprocessor computing systems.

\begin{figure}[h!]
    \centering
		\includegraphics[width=0.75\textwidth]{pictures/globalizerScheme.eps}
		\caption{The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)}
    \label{fig:globalizerScheme}
\end{figure}

\par
The Globalizer system architecture is presented in Figure \ref{fig:globalizerScheme}. The structural components are:
\begin{itemize}
  \item Block 0 consists of the procedures for computing the function values (criteria
  and constraints) for the optimization problem being solved; this is an external
  block with respect to the Globalizer system and is provided by the user per a predefined interface.
  \item Blocks 1-4 form the optimization subsystem and solve the global optimization
  problems (Block 1), nonlinear programming (Block 2), multicriterial optimization (Block 3),
  and general decision making problems (Block 4). It is worth noting the sequential
  scheme of interaction between these components --- the decision making problems are
  solved using the multicriterial optimization block, which, in turn, uses the nonlinear programming block, etc.
  \item Block 5 is a subsystem for accumulating and processing the search information;
  this is one of the basic subsystems --- the amount of search information for complex
  optimization problems may appear to be quite large on the one hand, but, on the other
  hand, the efficiency of the global optimization methods depends to a great extent on how
  completely all of the available search data is utilized.
  \item Block 6 contains the dimensional reduction procedures based on Peano evolvents;
  the optimization blocks solve the reduced (one-dimensional) optimization problems;
  Block 6 provides interaction between the optimization blocks and the initial
  multidimensional optimization problem.
  \item Block 7 organizes the choice of parallel computation schemes in the Globalizer
  system subject to the computing system architecture employed (the numbers of cores in the
  processors, the availability of shared and distributed memory, the availability of
  accelerators for computations, etc.) and the global optimization methods applied.
  \item Block 8 is responsible for managing the parallel processes when performing the
  global search (determining the optimal configuration of parallel processes, distributing
  the processes between computer elements, balancing computational loads, etc.).
  \item Block 9 is a common managing subsystem, which fully controls the whole
  computational process when solving global optimization problems.
  \item Block 10 is responsible for organizing the dialog interaction with users for
  stating the optimization problem, adjusting system parameters (if necessary), and
  visualizing and presenting the global search results.
  \item Block 11 is a set of tools for visualizing and presenting the global search results;
  the availability of tools for visually presenting the computational results enables
  the user to provide efficient control over the global optimization process.
\end{itemize}
%Figure 3. The Globalizer system architecture (Blocks 1-2, 5-7 have been implemented; Blocks 3-4 and 8-11 are under development)
\par
The Globalizer system comes in two different options:
\begin{itemize}
  \item Globalizer Pro is a research version of the system providing a full set of
  capabilities that are oriented toward solving of the most complicated global optimization problems.
  \item Globalizer Lite is a limited version of the Globalizer system accessible for free for
  solving global optimization problems with a medium level of difficulty, and is accessible free of charge.
\end{itemize}
\section{Results of Numerical Experiments}
\label{sec:experiments}
\subsection{Parallel Computations Using Intel Xeon Phi processors}
\subsection{Parallel Computations Using General Purpose Graphic Processors}
